### 2014-01-21 disc_arfi_mr.tex
integrated Raj's comments

We did not expect perfect agreement between MR and ARFI imaging volume or
tri-axial dimensions since each modality applies a different amount of external
compression to the rectal wall, which then deforms the prostate.  Pressure on
the prostate during ARFI imaging is modulated by the urologist when he locks
the transducer rotation apparatus in place prior to imaging, and deformation of
the prostate can be easily visualized during this alignment process.  It was
our goal to apply as little pressure while still maintaining uniform contact
throughout the image acquisition. This is critical since poor acoustic coupling
to the prostate through the rectal wall can greatly diminish ARFI image
displacement SNR.

>>>>

We did not expect perfect agreement between MR and ARFI imaging volume or
tri-axial dimensions since each modality applies a different amount of external
compression to the rectal wall, which then deforms the prostate.  The
endorectal coil places pressure on the prostate that is dependent on the
patient's anatomy and overall prostate size, while pressure on the prostate
during ultrasound imaging is modulated by the urologist when the transducer
rotation apparatus is locked into position prior to imaging.  The differential
deformation of the prostate can be easily visualized during the transducer
alignment process during ultrasound imaging.  It was our goal to apply as
little pressure while still maintaining uniform contact throughout the
ultrasound image acquisition. This is critical since poor acoustic coupling to
the prostate through the rectal wall can greatly diminish ARFI image
displacement SNR.

### 2014-01-21 intro.tex
integrated Raj's comments

\section{Introduction}\label{sect:intro}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer related
death.~\cite{Howlader2011} Screening with prostate specific antigen (PSA) and
digital rectal examination (DRE) has led to earlier PCa detection, but
performance of these measures is not optimal, leading to imprecise risk
assessment.  Additionally, the random nature of transrectal ultrasound
(TRUS)-guided biopsies can miss or underestimate the cancer
burden.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
deciding optimal management strategies for patients presenting with less
disease burden, whereas missed tumors on TRUS-guided biopsies can result in
inappropriate diagnoses.

The use of non-invasive prostate imaging could lead to improved PCa diagnosis,
risk-stratification, and management.  Magnetic resonance imaging (MRI) has been
available for use in the workup of patients with PCa since the early 1980s, but
early studies on its diagnostic accuracy were heterogeneous.  The more
recent ability to include functional parameters in PCa MRI analysis has yielded
promising results.~\cite{Gupta2013,Hricak2007} Among the MRI modalities
currently used in the study of PCa, it is well established that T2-Weighted
Imaging (T2WI) offers the best assessment of prostate anatomy based on its
ability to delineate prostatic margins, distinguish internal structures and
differentiate among the glandular zones.

@@

The goal of this study is to evaluate the ability of ARFI imaging to delineate
prostate zonal anatomy, specifically central gland and total prostate gland
\invivo as compared to endorectal MR T2WI.  Section~\ref{sect:background}
provides an overview of MR and ARFI imaging in the prostate and an overall
clinical motivation for prostate imaging.  Section~\ref{sect:methods} describes
the methods used to experimentally-acquire our imaging data, process the gross
pathological specimens post radical prostatectomy, and image process our
datasets.  The results of our analysis, including gross pathology and imaging
prostate axis and volume estimates, are presented in
Section~\ref{sect:results}, along with a statistical analysis of the bias and
variability associated with each imaging modalities measurements, all of which
is discussed in Section~\ref{sect:discussion}.

>>>>

\section{Introduction}\label{sect:intro}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer related
death.~\cite{Howlader2011} Approximately 1 in every 6 men will develop PCa
during their lifetime, with the median age of diagnosis at 67 years
old.~\cite{Howlader2011} PCa is also the second leading cause of cancer-related
death, with 1 in 36 men dying from the disease.  The National Cancer Institute
(NCI) estimates that 238,590 men will be diagnosed with PCa in 2013 and 29,720
will die from the disease.~\cite{Howlader2011}

PCa diagnosis usually begins by screening with prostate specific antigen (PSA)
and digital rectal examination (DRE).  More definitive diagnosis can be made by
random transrectal ultrasound (TRUS)-guided biopsies, which are then used to
provide the clinician with the proper Gleason score. The combination of these
factors, as well as staging, determines the appropriate therapy and prognosis.

PCa screening has led to earlier diagnosis of smaller tumors and more localized
disease; however, it is well known that the sensitivity and specificity of PSA
and DRE are not optimal. In addition, DRE has a low predictive value at lower
PSA ranges, and PSA yields many false
positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
risk of over-diagnosis and treatment of low-grade, and possibly clinically
insignificant, disease exists.  Moreover, due to the random nature of
TRUS-guided systematic biopsies, PCa located outside the routine sampling sites
can be missed and the extent of the cancer might be
underestimated.~\cite{Gupta2013,Cornud2012} For example, in a study by Mufarrij
\etal, 45.9--47.2\% of patients who were candidates for active surveillance,
but underwent radical prostatectomy, had a higher Gleason score on final
histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
may lead to inappropriate diagnosis, imprecise risk assessment and potentially
avoidable morbidity.

The use of non-invasive prostate imaging could lead to improved PCa diagnosis,
risk-stratification, and management.  Magnetic resonance imaging (MRI) has been
available for use in the workup of patients with PCa since the early 1980s, but
early studies on its diagnostic accuracy were heterogeneous.  The more
recent ability to include functional parameters in PCa MRI analysis has yielded
promising results.~\cite{Gupta2013,Hricak2007} Among the MRI sequences
currently used in the study of PCa, it is well established that T2-Weighted
Imaging (T2WI) offers the best assessment of prostate anatomy based on its
ability to delineate prostatic margins, distinguish internal structures and
differentiate among the glandular zones.

@@

The goal of this study is to evaluate the ability of ARFI imaging to delineate
prostate zonal anatomy, specifically central gland and total prostate gland
\invivo as compared to T2WI MR at 3T using an endorectal coil.
Section~\ref{sect:background} provides an overview of MR and ARFI imaging in
the prostate and an overall clinical motivation for prostate imaging.
Section~\ref{sect:methods} describes the methods used to experimentally-acquire
our imaging data, process the gross pathological specimens post radical
prostatectomy, and image process our datasets.  The results of our analysis,
including gross pathology and imaging prostate axis and volume estimates, are
presented in Section~\ref{sect:results}, along with a statistical analysis of
the bias and variability associated with each imaging modalities measurements,
all of which is discussed in Section~\ref{sect:discussion}.

### 2014-01-21 prostate_anatomy.tex
integrated Raj's comments

\subsection{Prostate Anatomy}
The prostate gland sits below the urinary bladder and surrounds the urethra.
Its superior borders include the bladder and seminal vesicles, and the
urogenital diaphragm delineates its inferior boundary. The gland is bordered
anteriorly by the pubic symphysis and posteriorly by the rectum.  The prostate
is separated from the rectum by a 2--3 mm fascial layer,~\cite{Jung2012} and it
can be easily palpated on rectal examination.

@@

Although not a true capsule, an outer band of fibromuscular tissue surrounds
the prostate.~\cite{Bonekamp2011} This ``capsule'' is important when assessing
the extraprostatic extension of cancer as tumor can spread by disrupting this
tissue. Two neurovascular bundles course posterior and lateral to the prostate,
which can also be invaded by malignant cells.  The prostate capsule is another
structure that was manually segmented in both MR and ARFI images in this study,
and its overall extent was used to delineate the total prostate gland.

>>>>

\subsection{Prostate Anatomy}
The prostate gland sits caudal to the urinary bladder and surrounds the urethra.
Its superior borders include the bladder and seminal vesicles, and the
urogenital diaphragm delineates its inferior boundary. The gland is bordered
anteriorly by the pubic symphysis and posteriorly by the rectum.  The prostate
is separated from the rectum by a 2--3 mm fascial layer,~\cite{Jung2012} and it
can be easily palpated on rectal examination.

@@

An outer band of fibromuscular tissue surrounds the anterior aspect of the
prostate.~\cite{Bonekamp2011} This boundary is important when assessing the
extraprostatic extension of cancer as tumor can spread by disrupting this
tissue. Two neurovascular bundles course posterior and lateral to the prostate,
which can also be invaded by malignant cells.  The prostate capsule is another
structure that was manually segmented in both MR and ultrasound images in this
study, and its overall extent was used to delineate the total prostate gland.
